---
title: "southernMaps package: mapping"
author: "Olivia Burge"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{southernMaps package: mapping}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

`r require(southernMaps)`


This package aims to provide Southern Hemisphere maps in easy forms for use. Therefore it provides functions, but also datasets. The datasets, for easy reference, are:

(1) nzSml, nzMed, nzHigh - shapefiles of the NZ coastline in NZTM projection
(2) nzSml84, nzMed84, nzHigh84 - shapefiles of the NZ coastline in WGS84 projection. 

If your data is in lat/lon, you want WGS84. The data load automatically, so you don't need to run anything. See below for examples

## NZ maps

The `map(nz)` command in base R returns different projection than expected (ie, not WGS84). If you want a map of NZ that's not too detailed, you can use `nzSml`. Note we use the `sf` package now!

```{r, fig.cap= "A low-res map of NZ in base map", fig.width = 8, fig.height = 6}
par(mfrow = c(1, 2))
plot(st_geometry(nzSml84))
plot(st_geometry(nzSml84), axes = TRUE) # with basic axes
par(mfrow = c(1,1))
```

If you want to plot it in **ggplot2**, you can use the `geom_sf` function in the `sf` package


```{r}


require(ggplot2)

ggplot() + 
  geom_sf(data = nzSml84) + 
  theme_minimal()
```

Minor amendments to remove the lat/long references and the gridlines:

```{r}

ggplot() + 
  geom_sf(data = nzSml84) + 
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank())

```


All of the above also applies to `nzMed` and `nzHigh`. 

## Creating a spatial points dataframe from a normal dataframe


```{r}
tp <- data.frame(
  Island = c("South Island", "North Island",
             "North Island", "Stewart Island", "South Island"),
  City = c("Christchurch", "Wellington", "Auckland", "Oban", "Dunedin"),
  long = c(172.63333,174.77557, 174.76667, 168.12736, 170.50361),
  lat = c(-43.53333, -41.28664, -36.86667, -46.89881, -45.87416))

wgs84 <- st_crs(4326) 
nztm <- st_crs(2193)

spatialCities <- spatial_df(data = tp, xcol = "long", ycol = "lat",
                            crs = wgs84)

cols <- c("forestgreen", "navyblue", "violetred4")
isl <- c("North Island", "South Island", "Stewart Island")

mapColours <- data.frame(cols = I(cols), isl)

ggplot() +
  geom_sf(data = nzSml, fill = "grey", col = NA) +
  geom_sf(data = spatialCities, aes(colour = Island),
          size = 2) +
  theme_minimal() + 
  scale_colour_manual(values = cols)


```

Next example - use the spatial object to get distances between cities

```{r}


# example - transform to NZTM, 
# get distance between cities
# use the units package to convert to km

citiesNZTM <- st_transform(spatialCities,
                              crs = 2193)

sf::st_distance(citiesNZTM, byid = TRUE) %>%
  units::set_units(., "km")

```

## Plotting small islands in high resolution

This section also relates to plotting small areas of NZ in high detail. The key function is `find_my_island()`, which returns islands from the data object `detailed_nz_islands`. It returns the islands its found by searching across the various shapefile fields. Easiest to show by example:


```{r}
isle <- find_my_island("codfish", proj = "nztm") #detailed_nz_islands %>% filter(grepl("codfish", tolower(name)))#
plot(isle)
```

And to plot in **ggplot2**:

```{r}

idf <- ggplot() +
  geom_sf(data = isle, colour = "grey",
          fill = "white") +
  geom_polygon(colour = "grey", fill = "white")+
  labs(x = "Longitude E", y = "Latitude S", 
       title = "Codfish Island") 

idf
```

You can also do multiple islands at once: 

```{r}
# multiple islands
many <- find_my_island(c("auckland", "campbell"))#detailed_nz_islands %>% filter(grepl("auckland|campbell", tolower(grp_name)) |
                                        # grepl("auckland|campbell", tolower(name))) #find_my_island(c("auckland", "campbell"))



ggplot() +
  geom_sf(data = many) +
  labs(title = "Offshore islands", subtitle = "Auckland and Campbell islands") + 
  theme_void() # no axes, nothing 
```

## Adding points

This has nothing to do with the southernMaps package, but for completeness, we use the `idf` plot from above and add to it. Firstly we create some "points", then add them. You will already have a dataframe it is assumed with the points in. Make sure it's in the same projection as the map! You can check a spatial file's projection like so: `st_crs(isle)`. you can't do this with dataframes like `isle_df`. 

```{r, fig.width = 8, fig.height = 7}

some_points <- st_sample(isle, size= 8) %>%
  st_as_sf(.) %>%
  mutate(type = c(rep("Control", 4), rep("Treatment", 4)),
         siteName = c("Umbrella", "Peach", "Car", "Dog", 
                      "Film", "Laptop", "Backpack", "Plate"))


idf +
  geom_sf(data = some_points, aes(fill = type),
          shape = 21, colour = "black") +
  scale_fill_manual(values = c("black", "white")) +
  theme(legend.position = "bottom") +
  labs(fill = "Treatment") +
  geom_sf_text(data = some_points %>% filter(siteName != "Dog") %>% 
                 filter(siteName != "Peach"), 
          aes(label = siteName),
            nudge_x = 50, 
            hjust = 0) +
    geom_sf_text(data = some_points %>% filter(siteName == "Dog"), 
            aes(label = siteName),
            nudge_x = 70,
            nudge_y = -100,
            hjust = 0) +
      geom_sf_text(data = some_points %>% filter(siteName == "Peach"), 
            aes(label = siteName),
            nudge_x = 70,
            nudge_y = 100,
            hjust = 0)
```

